Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk

### 2021-09-28 VegOzone.tex
Restructure and include remark on grassland CL

We define two different hypothetical temperature acclimations: subarctic and cold. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to default values. In the same way, \emph{subarctic} constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. In other words, we aim to increase the overlapping area between the $f_\mathrm{temp}$ and the temperature PDF of the GS. The most extreme of the two acclimation types is \emph{subarctic}.
In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from five months GS in two climatological periods. The periods 1990s and 2000s have been chosen to indicate that the vegetation is subject to ongoing climate change Thus, the predominant acclimation at present is not known.

@@

The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here. Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when relying on generic MM parameterizations. We will look at the specific PFTs in more detail in the following.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light parameterization is pronounced. With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses. However, taking the later start and overall shorter GS in 2018 into account (bespoke GS), this is reversed for the \emph{subarctic} acclimation type. This can be explained by favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke growing season. The temperature acclimation-dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the beginning and therefore the least biased of the generic MM parameterizations. Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions. The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities is medium and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large response to the temperature parameterizations as deciduous trees. Again, we find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{subarctic}) for each bespoke PFT. The estimates are listed in Table~\ref{tab:biomass_reduction}. The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on. For deciduous trees and perennial grassland, total biomass was reduced in comparison with the default generic MM parameterizations for both years. \textbf{TODO: Discussion about CL for grassland!} Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties. This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening. The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.

>>>>

We define two different hypothetical temperature acclimations: subarctic and cold. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to default values. In the same way, \emph{subarctic} constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF of the GS. The most extreme of the two acclimation types is \emph{subarctic}.
In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from five months GS in two climatological periods. The periods 1990s and 2000s have been chosen to indicate that the vegetation is subject to ongoing climate change Thus, the predominant acclimation at present is not known.

@@

The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here. Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.

For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light parameterization is pronounced. With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses. However, taking the bespoke GS into account, this is reversed for the \emph{subarctic} acclimation type. This can be explained by more favorable growing conditions in 2019. Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS. The temperature acclimation-dependent difference in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the start and therefore differs least. Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions. The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large response to the temperature parameterizations as deciduous trees. Again, we find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization (\emph{subarctic}) for each bespoke species or PFT. The estimates are listed in Table~\ref{tab:biomass_reduction}. The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on. Total biomass of deciduous trees and perennial grassland was more strongly reduced in comparison with the default MM parameterizations for both years.

The CL for perennial grassland is placed higher than for the other species. From an agro-economical point of view in our focus area, a loss of biomass of more than $10\,\unit{\%}$ can be more severe compared to more productive areas in central Europe and may need adjustment.

Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties. This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening. The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.
